% What Does Precedent Retrieval Actually Retrieve?
% Target: arXiv cs.SE preprint (cross-list cs.IR), later MSR technical track.
% Build: latexmk -pdf main.tex  (or upload the paper/ directory to Overleaf).
% v0.3 — reviewer pass: main finding surfaced, figures added, deployment A/B
% section added (eval/ab/RESULTS.md), system section expanded, abstract cut.
% Every number traces to docs/EVAL.md or eval/ab/RESULTS.md.
\documentclass[sigconf,nonacm]{acmart}

\usepackage{booktabs}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning}

\begin{document}

\title{Ranking Is Not the Bottleneck: A Leak-Free Evaluation of
  Precedent Retrieval over Merged Pull-Request History}

\author{Vladimir Ivanov}
\affiliation{%
  \institution{Independent Researcher}
  \country{Georgia}}
\email{vladimir@ivanovresearch.dev}

\begin{abstract}
Precedent retrieval serves an AI coding agent the merged pull-request
history of its own repository, each item a problem--solution pair. We ask
what its evaluation actually measures. On a backport-deduplicated corpus of
575 merged PRs (\texttt{dotnet/efcore}), known-item retrieval is strong
(recall@5 = 0.90 over 218 held-out queries) --- yet three further
instruments show that ranking quality is not the binding constraint.
(1)~In a leak-free temporal holdout, a metadata-linked relevance label
finds only 5 of 278 new tasks answerable, while blind human judgment finds
29 of 30: \textbf{the standard label misses 93--97\% of true precedents}.
(2)~Under TREC-style pooling, \textbf{29\% of judged-relevant precedents
surface only outside the primary embedding retriever} --- single-system
evaluation is structurally biased. (3)~An LLM judge fails its
pre-registered calibration gate (Cohen's $\kappa = 0.39 < 0.6$; recall
0.35 against human labels), missing precisely the convention-transfer
cases pooling exists to capture; it is retained advisory-only. (4)~A
12-task paired deployment probe shows no outcome uplift on a shallow
3.5-month corpus but also zero harm; the only two value events were
codebase-convention transfers from the deepest history available.
Together these results locate the bottleneck of precedent retrieval in
\emph{relevance definition and corpus depth}, not ranking --- and imply
that retrieval evaluations which stop at known-item metrics measure the
wrong thing.
\end{abstract}

\keywords{mining software repositories, retrieval evaluation, AI coding
agents, temporal holdout, TREC pooling, LLM-as-judge calibration}

\maketitle

% ---------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}

AI coding agents fail most where codebases are oldest: the conventions,
workarounds, and edge-case lore of a mature repository live not in its
current source --- which context tools already embed --- but in years of
merged pull requests that no base model has seen. \emph{Precedent
retrieval} targets exactly that gap: index every merged PR as a
problem--solution pair (the issue text that states what was wrong, and the
diff the team actually accepted), and let an agent working on a new task
retrieve how similar problems were actually solved \emph{in this codebase}
(Figure~\ref{fig:pipeline}).

The uncomfortable question is how to know whether such a system works.
The standard answer --- known-item retrieval, ``does the PR known to fix
this issue rank in the top-$k$?'' --- is cheap, reproducible, and, we will
argue, systematically misleading when read as a statement about deployment
utility. It measures ranking quality. It says nothing about whether a
\emph{genuinely new} task can find a relevant precedent that predates it,
which is the question that decides whether precedent retrieval is useful
at all.

We built the instruments to answer that harder question over the
\texttt{dotnet/efcore} history, holding a strict discipline: gates
pre-registered before results existed, relevance labeled blind, leakage
controls checked rather than assumed. The resulting measurements ---
including two negative results we report at full strength --- compose into
one finding:

\begin{quote}
\textbf{Ranking is not the bottleneck of precedent retrieval. The
relevance definition and the depth of the indexed history are.}
\end{quote}

Specifically:

\begin{itemize}
  \item[\textbf{F1.}] Ranking quality is high and stable: recall@5 = 0.90
    on a backport-deduplicated known-item benchmark
    (\S\ref{sec:knownitem}). Yet the same system, evaluated temporally
    with the field's default relevance label (explicit issue links),
    appears almost useless: 5 of 278 new tasks answerable. Blind human
    judgment on a 30-task probe reverses the verdict --- 29 of 30 tasks
    have a genuine prior precedent --- exposing that \textbf{the
    metadata label misses 93--97\% of true precedents}
    (\S\ref{sec:power}).
  \item[\textbf{F2.}] Relevance, once judged rather than inferred, is
    \emph{plural}: 29\% of judged-relevant precedents surface only
    through retrieval legs other than the primary embedding system
    (\S\ref{sec:pooling}). Evaluating a retriever against candidates it
    alone surfaces is structurally self-confirming.
  \item[\textbf{F3.}] The obvious scaling move --- an LLM judge ---
    fails its pre-registered calibration gate ($\kappa \le 0.39 < 0.6$),
    and its misses concentrate on exactly the convention-transfer
    relevance that pooling exists to capture (\S\ref{sec:judge}). We keep
    it advisory-only.
  \item[\textbf{F4.}] In a 12-task paired deployment probe with live
    agents, a shallow corpus (3.5 months of history) produced no outcome
    uplift --- and no harm: agents respected the system's
    below-threshold signal in 8 of 12 tasks, zero runs were misled, and
    the only two value events were convention transfers from the deepest
    history available (\S\ref{sec:ab}). Utility tracks corpus depth, not
    ranking headroom.
\end{itemize}

Our contributions: (i)~the first leak-free temporal-holdout protocol for
retrieval over PR history, with leakage controls that are checked, not
assumed; (ii)~empirical evidence that metadata-linked relevance labels
underpower temporal instruments by an order of magnitude, diagnosed by a
blind labeled probe; (iii)~a TREC-pooled, human-calibrated relevance
definition for PR-history retrieval, validated per retrieval leg; and
(iv)~a measured negative result on LLM-as-judge labeling for this task,
with its failure mode characterized. The system, harness, and hash-pinned
manifests are open source.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  node distance=3.5mm,
  box/.style={draw, rounded corners=1pt, minimum height=5.5mm,
    minimum width=21mm, font=\small, align=center},
  lbl/.style={font=\scriptsize\itshape, text=black!60},
  arr/.style={-{Stealth[length=2mm]}, thick}]
  \node[box] (pr) {merged PR};
  \node[box, below left=4mm and -8mm of pr] (prob) {problem\\ \scriptsize title + body};
  \node[box, below right=4mm and -8mm of pr] (sol) {solution\\ \scriptsize diff};
  \node[box, below=13mm of pr, minimum width=44mm] (db)
    {PostgreSQL + pgvector\\ \scriptsize two embeddings per PR, HNSW};
  \node[box, below=4mm of db, minimum width=44mm] (mcp)
    {retrieval server (MCP)\\ \scriptsize weighted composite + threshold};
  \node[box, below=4mm of mcp] (agent) {AI coding agent};
  \draw[arr] (pr) -- (prob);
  \draw[arr] (pr) -- (sol);
  \draw[arr] (prob) -- (db) node[midway, left, lbl] {embed};
  \draw[arr] (sol) -- (db) node[midway, right, lbl] {embed};
  \draw[arr] (db) -- (mcp);
  \draw[arr] (mcp) -- (agent) node[midway, right, lbl] {ranked precedents};
\end{tikzpicture}
\caption{Precedent retrieval. Each merged PR is indexed as a
problem--solution pair; an agent queries in natural language and receives
ranked precedents with a confidence label, or an explicit below-threshold
signal.}
\label{fig:pipeline}
\end{figure}

% ---------------------------------------------------------------------------
\section{Background and Related Work}
\label{sec:related}

\paragraph{Known-item evaluation.}
Known-item retrieval --- find the one document known to answer the query ---
is among the oldest evaluation designs in IR~\cite{sparck1975ideal}. It
requires no relevance judging beyond the query--target link, which makes it
cheap and reproducible, but it measures only whether a target the system is
\emph{told} exists can be ranked highly. Voorhees's work on judgment
variation~\cite{voorhees2000variations} and Zobel's analysis of pooling
reliability~\cite{zobel1998reliable} establish the point we build on: what
one measures is determined by how relevance is defined, and narrow
definitions bias conclusions.

\paragraph{Temporal splits and leakage.}
Time-respecting evaluation is standard advice in mining software
repositories, yet leakage enters through subtle channels: retrospectively
edited artifacts, re-ingested corpora that silently move the split
boundary, and labels computed over post-cutoff state. Kalliamvakou et
al.~\cite{kalliamvakou2014promises} catalogue the perils of mining GitHub
data at face value; our design (\S\ref{sec:temporal}) treats their
warnings as checklist items --- e.g., PR descriptions are mutable, so the
split must freeze on merge timestamps and original ingest state, not on
current body content. Benchmarks for repository-level agent tasks such as
SWE-bench~\cite{jimenez2024swebench} face analogous contamination
questions from model training data; our instrument addresses the
retrieval-side analogue.

\paragraph{Pooling.}
Judging every (task, candidate) pair is infeasible at corpus scale. TREC
pooling~\cite{voorhees2000variations,zobel1998reliable} unions the top
candidates of multiple independent systems into a judgment pool; relevance
is then judged only inside the pool. Pool composition is a known source of
bias --- a pool built from one system family under-represents what only
other families retrieve. Our pool (\S\ref{sec:pooling}) unions an
embedding retriever, a lexical BM25 leg~\cite{robertson2009bm25}, and a
second-family embedder (bge-m3~\cite{chen2024bge}), and we validate each
leg's contribution against blind human labels.

\paragraph{LLM-as-judge.}
Scaling relevance labeling with an LLM judge is attractive and increasingly
common~\cite{zheng2023judging}. The known risk is that judge agreement with
humans is task-dependent and must be measured, not assumed. We
pre-registered a Cohen's $\kappa$~\cite{cohen1960} authority gate and
report a failure against it (\S\ref{sec:judge}); the failure mode ---
misses concentrated on convention-transfer relevance --- is, to our
knowledge, underdocumented and directly actionable for others building
retrieval evaluations over software history.

\paragraph{Context for coding agents.}
Contemporary context tools embed \emph{current} code (semantic code
search~\cite{husain2019codesearchnet}, repository indexing in commercial
assistants). Precedent retrieval is complementary: it embeds how the code
has historically been \emph{changed} --- problem--diff pairs from merged
PRs~\cite{gousios2014exploratory} --- which is precisely the knowledge that
is absent from current source and from base-model training in private,
legacy codebases.

% ---------------------------------------------------------------------------
\section{System Under Evaluation}
\label{sec:system}

\textsc{senrah} is an open-source, self-hosted retrieval server exposed to
agents over the Model Context Protocol~\cite{mcp2024}. We describe the
pipeline in the detail needed to interpret the experiments.

\paragraph{What is indexed.}
Ingestion pulls every merged PR (resumable traversal; bots and automation
filtered by stop-list and title patterns; oversized PRs excluded). Two
texts are constructed per PR: the \emph{problem} text --- PR title
concatenated with the PR body --- and the \emph{solution} text --- the raw
unified diff. Each is truncated by \emph{model tokens}, never characters,
with independent budgets (1{,}500 tokens for problem, 6{,}000 for diff),
so titles always survive truncation.

\paragraph{Embedding and storage.}
Both texts are embedded independently with
\texttt{text-embedding-3-small}~\cite{openai2024embeddings} (1{,}536
dimensions) and stored in PostgreSQL with pgvector; each embedding column
carries an HNSW index~\cite{malkov2020hnsw} (cosine distance), so
incremental ingest requires no index rebuild. The embedding model and
version are recorded per row, giving a re-embedding migration path.

\paragraph{Scoring.}
A query is embedded once. Approximate nearest-neighbor search oversamples
candidates on \emph{both} columns, and candidates are re-ranked by a
weighted composite of the two cosine similarities:
$\mathit{score} = w_p \cdot \mathit{sim}_\mathit{problem} + w_s \cdot
\mathit{sim}_\mathit{solution}$.
The shipped default is $w_p/w_s = 0.6/0.4$; the frozen evaluation protocol
uses $0.7/0.3$ with score threshold $0.45$, fixed when the protocol was
registered. (A systematic weight ablation is future work; the deployment
probe in \S\ref{sec:ab} suggests weight tuning is second-order relative to
corpus depth.) Results below the threshold are withheld and replaced by an
explicit below-threshold signal --- a design choice that turns out to
matter for deployment safety (\S\ref{sec:ab}).

% ---------------------------------------------------------------------------
\section{Instrument 1: Known-Item Retrieval}
\label{sec:knownitem}

\paragraph{Definition.}
For a known issue whose fixing PR is known, does that PR appear in the
top-$k$ results when the issue text is used as the query?

\paragraph{Corpus and protocol.}
\texttt{dotnet/efcore}, deduplicated: 575 merged PRs (merged 2024-04-06
through 2026-06-12), 218 held-out queries. The manifest is hash-pinned and
the run is deterministic.

\paragraph{Deduplication is load-bearing.}
A naive corpus double-counts backports and cherry-picks: the ``same'' fix
appears several times, so \emph{a} hit is easy and recall inflates. We
detect backport clusters from stored diffs using union-find over
\emph{corroborated} edges --- diff similarity alone is never allowed to
merge two PRs --- and score \textbf{per-cluster}: a hit on any cluster
member counts once, and distractors are likewise counted per-cluster.

\begin{table}[h]
  \caption{Known-item retrieval on deduplicated \texttt{dotnet/efcore}
  (575 merged PRs, 218 held-out queries, per-cluster scoring).}
  \label{tab:knownitem}
  \begin{tabular}{lr}
    \toprule
    Metric & Value \\
    \midrule
    recall@1 & 0.71 \\
    recall@5 & 0.90 \\
    recall@10 & 0.93 \\
    MRR@10 & 0.79 \\
    \bottomrule
  \end{tabular}
\end{table}

\paragraph{Why this is not the whole story.}
Known-item recall answers ``can the system rank a target it is told
exists.'' It says nothing about coverage for genuinely new work. Worse, on
a deeper corpus known-item recall can \emph{fall}, because more PRs means
more near-duplicate distractors competing for top-$k$ slots --- so
optimizing this number alone penalizes exactly the corpus growth that
\S\ref{sec:ab} will show is where utility lives.

% ---------------------------------------------------------------------------
\section{Instrument 2: Leak-Free Temporal Holdout}
\label{sec:temporal}

\begin{figure}[t]
\centering
\begin{tikzpicture}[font=\small]
  \draw[thick, -{Stealth[length=2.5mm]}] (0,0) -- (8.2,0)
    node[below=1mm, pos=0.98, font=\scriptsize] {time};
  \draw[very thick, red!70] (5.2,-0.28) -- (5.2,0.75)
    node[above, font=\small] {cutoff $T$};
  \draw[decorate, thick] (0.1,0.12) -- (5.1,0.12);
  \node[above, font=\small] at (2.6,0.3)
    {retrievable corpus \scriptsize($\mathit{merged\_at} < T$)};
  \draw[decorate, thick] (5.3,0.12) -- (8.0,0.12);
  \node[above, font=\small] at (6.7,0.3)
    {new tasks \scriptsize($> T$)};
  \foreach \x in {0.5,1.1,1.6,2.4,3.0,3.7,4.3,4.8}
    \node[circle, fill=black!55, inner sep=1.1pt] at (\x,0) {};
  \foreach \x in {5.6,6.3,7.0,7.6}
    \node[circle, draw=black!70, inner sep=1.1pt] at (\x,0) {};
  \draw[-{Stealth[length=2mm]}, dashed, black!60]
    (6.3,-0.12) to[bend left=25] (3.0,-0.14);
  \node[below, font=\scriptsize, black!60] at (4.6,-0.5)
    {does a relevant precedent strictly before $T$ get retrieved?};
\end{tikzpicture}
\caption{Temporal holdout. The split freezes on merge timestamps and
original ingest state; labels freeze together with the corpus boundary;
the scorer applies the product's own score threshold.}
\label{fig:temporal}
\end{figure}

\paragraph{Design.}
Pick a cutoff time $T$. Everything merged strictly before $T$ is the
retrievable corpus; everything merged strictly after $T$ is the incoming
task stream (Figure~\ref{fig:temporal}). For each post-$T$ task: does a
relevant precedent exist strictly before $T$, and does the system retrieve
it? This simulates the deployed setting --- an agent working today,
retrieving from history.

\paragraph{Leakage controls (where temporal evals die).}
\begin{itemize}
  \item The corpus filter is strictly $\mathit{merged\_at} < T$; queries
    strictly $\mathit{merged\_at} > T$. The split freezes on
    $\mathit{merged\_at}$ and the \emph{original} ingest timestamps ---
    never on current PR-body state, which is editable after the
    fact~\cite{kalliamvakou2014promises}.
  \item Relevance labels and the corpus boundary freeze together, so a
    later re-ingest cannot silently move the boundary under the labels.
  \item The scorer applies the product's own below-threshold cutoff, so
    the instrument measures what a user actually receives, not an
    idealized ranking.
\end{itemize}

\paragraph{Materialization.}
The full multi-year history was ingested for this instrument
(\texttt{efcore}: 487 $\rightarrow$ 8{,}449 PRs; one ingest, one index,
every depth rung materializable from it), and a deep cluster map built
(9{,}594 PRs, 397 multi-member backport clusters, hash-pinned).

% ---------------------------------------------------------------------------
\section{The Power Failure and Its Diagnosis}
\label{sec:power}

With relevance defined the way the field usually defaults to --- a
post-$T$ task is \emph{answerable} iff a metadata-linked prior PR exists
strictly before $T$ --- the harness came back underpowered
(Table~\ref{tab:underpowered}).

\begin{table}[h]
  \caption{Answerable post-$T$ tasks under the metadata-linked relevance
  label, against a pre-registered power floor of $\ge 80$.}
  \label{tab:underpowered}
  \begin{tabular}{rrr}
    \toprule
    Cutoff $T$ (days back) & Post-$T$ tasks w/ issue & Answerable \\
    \midrule
    365 & 278 & \textbf{5} \\
    455 & 306 & 4 \\
    545 & 345 & 4 \\
    \bottomrule
  \end{tabular}
\end{table}

Five is not a measurement; it is noise. The intellectually cheap move is
to pivot the evaluation to the known-item number that already looks good
and never mention the failure. We did not, because $n{=}5$ has two
\emph{opposite} explanations with opposite consequences: either
(\textbf{A}) the label is too strict --- metadata-linked relevance misses
precedents transferred by convention rather than by explicit issue link
--- or (\textbf{B}) precedents are genuinely rare and the honest output is
a negative coverage result.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  node distance=2.5mm and 8mm,
  box/.style={draw, rounded corners=1pt, minimum height=5.5mm,
    font=\scriptsize, align=center, inner sep=2.5pt},
  verdict/.style={font=\scriptsize\bfseries},
  arr/.style={-{Stealth[length=1.8mm]}, thick}]
  \node[box, minimum width=34mm] (task)
    {new task \#38297\\ Cosmos: translate top-level \texttt{Any()}};
  \node[box, below left=6mm and -14mm of task, minimum width=30mm] (meta)
    {metadata label:\\ no linked prior PR};
  \node[box, below right=6mm and -14mm of task, minimum width=34mm] (human)
    {human judgment:\\ precedent \#37603 --- same\\
     \texttt{WithIn} root-query pattern};
  \node[verdict, below=2mm of meta, text=red!70] {answerable: NO};
  \node[verdict, below=2mm of human, text=green!50!black] {answerable: YES};
  \draw[arr] (task) -- (meta);
  \draw[arr] (task) -- (human);
\end{tikzpicture}
\caption{Why metadata labels undercount (real case from the deployment
probe): the fix for \#38297 reused the root-query detection pattern
introduced by \#37603 --- a genuine precedent transferred by convention,
invisible to issue-link metadata.}
\label{fig:metadata}
\end{figure}

One cannot tell A from B by staring at the number. We drew 30 random
post-$T(365)$ tasks and hand-checked, independently of metadata links,
whether a genuine relevant precedent existed strictly before $T$. A
genuine pre-$T$ precedent existed for at least 14 of 30 under a strict
reading, and up to roughly 29 of 30 under a lenient one: \textbf{the
metadata label undercounted true relevance by 14--29$\times$ ---
equivalently, it missed 93--97\% of true precedents}
(Figure~\ref{fig:metadata} shows a representative real case). Verdict:
\textbf{A}. The first number was not the finding --- it was a symptom, and
the probe identified the disease.

% ---------------------------------------------------------------------------
\section{Rebuilding Relevance: TREC Pooling with Human Calibration}
\label{sec:pooling}

If metadata links undercount relevance, relevance must be judged, not
inferred. Judging every pair is infeasible, so we pool: for each task, the
top candidates of three independent retrieval legs are unioned into a
judgment pool --- \texttt{eval} (the evaluated production embedding
retriever), \texttt{bm25\_unique} (candidates only lexical
BM25~\cite{robertson2009bm25} surfaces), and \texttt{bge\_unique}
(candidates only a second-family embedder, bge-m3~\cite{chen2024bge},
surfaces). Each leg is included because each contributes relevant
precedents the others miss; a BM25-only pool recalled just 72\% of
judged-relevant candidates.

\paragraph{Blind gold set.}
We hand-labeled 270 (task, candidate) pairs --- 30 tasks $\times$ 9
candidates, three per leg --- \emph{before} looking at any depth result.
Density: 65 yes / 205 no (24\% positive). Coverage resolves the power
failure: \textbf{29 of 30} tasks are answerable under judged relevance
(versus 5 of 278 under the metadata label).

\begin{table}[h]
  \caption{Per-leg precision on the blind gold set.}
  \label{tab:legs}
  \begin{tabular}{lrr}
    \toprule
    Leg & relevant / total & precision \\
    \midrule
    \texttt{eval} & 46 / 90 & 0.51 \\
    \texttt{bge\_unique} & 11 / 90 & 0.12 \\
    \texttt{bm25\_unique} & 8 / 90 & 0.09 \\
    \bottomrule
  \end{tabular}
\end{table}

The production leg is precise --- every second candidate is relevant. The
unique legs are low-precision but decisively non-zero: together they
contribute \textbf{19 of 65 relevant precedents (29\%) that the main leg
missed} --- the classical pooling trade of coverage bought at the cost of
precision~\cite{zobel1998reliable}. Neither the lexical nor the
second-family leg can be dropped without biasing the pool: an evaluation
that judges only what the evaluated system retrieves is structurally
self-confirming.

% ---------------------------------------------------------------------------
\section{A Negative Result: The LLM Judge Fails Its Gate}
\label{sec:judge}

Hand-labeling 30 tasks rescued statistical power, but the instrument needs
80+ tasks to clear its pre-registered floor, and hand-labeling roughly 720
pairs per corpus does not scale. The plan was an LLM judge to label the
full pool --- \emph{conditional} on passing a pre-registered authority
gate: Cohen's $\kappa$ (judge vs.\ human) $\ge 0.6$~\cite{cohen1960}.

It did not pass (Table~\ref{tab:judge}).

\begin{table}[h]
  \caption{LLM-judge calibration against human gold labels. The
  pre-registered authority gate was $\kappa \ge 0.6$.}
  \label{tab:judge}
  \begin{tabular}{lrl}
    \toprule
    Judge (framing) & $\kappa$ vs.\ human & Verdict \\
    \midrule
    Sonnet 4.6 (phase-9 gold) & 0.27 & below floor \\
    Opus 4.8 (phase-9 gold) & 0.39 & below floor \\
    Sonnet 4.6 (temporal gold) & \textbf{0.39} & \textbf{below floor} \\
    \bottomrule
  \end{tabular}
\end{table}

On the temporal gold set the confusion matrix (human $\times$ judge) is
TP = 23, FN = 42, FP = 7, TN = 198 --- precision 0.77, recall
\textbf{0.35}. The judge misses 65\% of human-relevant precedents, and the
misses concentrate exactly where it hurts: 10 of 11 \texttt{bge\_unique}
and 6 of 8 \texttt{bm25\_unique} relevant precedents --- the semantic and
convention-transfer cases the pool was widened to capture. The judge's
blind spot is aligned with the pool's reason for existing.

\paragraph{Decision.}
The judge is retained \textbf{advisory-only}: its yes-labels are reasonably
trustworthy (precision 0.77); its no-labels are not (recall 0.35), and it
can never override the recall guardrail. We prefer reporting a slower,
human-anchored result over laundering a coverage number through a labeler
we have measured to be unreliable on the cases that matter. Consequently,
the full-power forward-coverage number remains \emph{open} and
human-labeling-bound; we track it as such rather than substituting a
pretend value.

% ---------------------------------------------------------------------------
\section{Deployment Probe: A Paired A/B on Live Tasks}
\label{sec:ab}

Retrieval metrics, however honest, are proxies. As a reality check we ran
a paired A/B on 12 real \texttt{efcore} tasks (well-specified bug reports,
selected by a deterministic filter): the same agent model solved each task
twice --- once with precedent retrieval available (treatment), once
without (control) --- under a protocol frozen before the last pairs
landed. Outcome metrics: file-level and symbol-level recall of the
agent's patch against the actual merged fix. The corpus available to the
treatment arm was shallow: 264 PRs, roughly 3.5 months of history.

\paragraph{Outcome.}
A dead heat: 6 control / 4 treatment / 2 ties (two-sided sign test on the
10 non-ties, $p \approx 0.75$); mean file recall 0.55 vs.\ 0.54. On these
outcome metrics the system neither helped nor hurt.

\paragraph{What the outcome metrics do not show.}
Categorizing all 12 treatment runs:
\begin{itemize}
  \item \emph{Exact solution shape retrieved:} 0.
  \item \emph{Codebase-convention transfer:} 2 clear (+2 weak). The fix
    for \#38297 reused the root-query detection pattern of \#37603; the
    fix for \#38271 adopted \#37652's pattern of excluding
    full-text/vector/JSON indexes from a DDL transform. Both transfers
    came from the deepest history present in the shallow corpus, and both
    pairs also scored at or near the top of the outcome metrics.
  \item \emph{Found but useless:} 8 --- scores clustered 0.31--0.49,
    mostly below the 0.45 threshold; agents read the below-threshold
    signal and ignored the results.
  \item \emph{Misled by a precedent:} \textbf{0.} The explicit
    below-threshold signal did its protective job.
\end{itemize}

\paragraph{Reading.}
On a 3.5-month corpus, genuine precedents for novel bugs are rare --- the
correct answer to most queries is ``nothing relevant,'' which the system
reports honestly and agents respect. The two value events that did occur
were convention transfers from the oldest history available, and 9 of 12
pairs converged on essentially the same fix regardless of arm ---
consistent with the task profile (well-specified bug reports that already
pin the fix location). This locates the improvement axis in \emph{corpus
depth and task profile}, not in re-ranking the same shallow corpus ---
converging, from the deployment side, with what the temporal instrument
showed from the evaluation side.

% ---------------------------------------------------------------------------
\section{Discussion}
\label{sec:discussion}

\paragraph{The composite finding.}
Four instruments, one picture. Known-item says ranking is strong (0.90).
The temporal instrument says the field's default relevance label hides
93--97\% of true precedents. Pooling says 29\% of true relevance is
invisible to any single-system evaluation. The deployment probe says
utility on a shallow corpus is bounded by content, not ranking --- and
that an honest below-threshold signal converts ``nothing relevant'' from
a failure mode into safe behavior. None of these facts is visible from
the known-item number alone; three of the four would be invisible even in
principle.

\paragraph{Why merged-PR history?}
Among the candidate substrates for precedent retrieval, merged PRs are the
only one that bundles a stated problem, a reviewed and \emph{accepted}
solution, and a timestamp of acceptance. Commits are too granular and
their messages rarely state the problem; issues state problems without
solutions; Stack Overflow answers lack codebase specificity and licensing
clarity; current-code search~\cite{husain2019codesearchnet} shows what the
code \emph{is}, not how the team changes it. The merged PR is where a
team's intent, review standards, and actual practice
coincide~\cite{gousios2014exploratory}.

\paragraph{Design guidance.}
Four practices did disproportionate work. (1)~\emph{Pre-register the
gates}: the $\kappa \ge 0.6$ authority floor and the $\ge 80$-task power
floor were fixed before any result existed, so a disappointing result
could not be redefined into a passing one. (2)~\emph{Blind-label before
looking}. (3)~\emph{Treat the first number as a symptom}: $n{=}5$ was not
the answer; the probe that explained it was. (4)~\emph{Report negative
results as results}: a judge that fails calibration and a coverage
question that remains open are stated plainly, with the numbers.

\paragraph{Threats to validity.}
Everything here is validated on a single corpus (\texttt{dotnet/efcore});
efcore's disciplined issue-linking culture likely makes the metadata-label
finding \emph{conservative} for messier repositories. The production
retriever uses a single embedding family; the pool mitigates but cannot
eliminate family bias. The deployment probe is small ($N{=}12$), uses one
agent model, and over-represents well-specified bug reports --- the task
profile least likely to benefit from precedents. Judge prompting was
varied across two framings but not exhaustively engineered; a materially
better judge may yet pass the gate, which is why the gate is
pre-registered rather than abandoned. A systematic ablation of the
problem/solution weights, and an error taxonomy over the known-item
misses, are planned follow-ups. The corpus is English-language.

% ---------------------------------------------------------------------------
\section{Conclusion}
\label{sec:conclusion}

Precedent retrieval over merged-PR history is easy to demonstrate and hard
to measure. Measured honestly, its ranking is strong (recall@5 = 0.90),
its default relevance labels are an order of magnitude too narrow
(93--97\% of true precedents missed), 29\% of true relevance is invisible
to single-system evaluation, the obvious scalable labeler fails its
calibration gate ($\kappa = 0.39 < 0.6$) precisely on the cases that
matter, and deployed utility tracks corpus depth rather than ranking
headroom. The instruments stand; the shortcuts do not. We release the
system, the harness, and the hash-pinned manifests, and we report the open
questions as open. We hope future benchmarks for retrieval in software
engineering adopt leak-free temporal evaluation with judged relevance as
the default --- and treat known-item retrieval as what it is: a necessary
smoke test, not a verdict.

% ---------------------------------------------------------------------------
\section*{Data Availability}
The system, evaluation harness, gold-set labels, and hash-pinned manifests
are available in the open-source repository:
\url{https://github.com/ivanovresearch/senrah} (evaluation account in
\texttt{docs/EVAL.md}; deployment-probe protocol and per-run scores in
\texttt{eval/ab/}).

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}
